Space transformation in Multidimensional IRT (Theorem)

### 2022-07-27 notebooks/Space_transformation_theorem.qmd
Cosine director lemmas continued [WIP]

Let's consider a dichotomous item $`r ITEM_INDEX`$ such that the probability
that respondent with trait $`r LS_BASIS_EQ`$ gives a correct[^1] response is
given by

@@

Let $`r BASIS_CHANGE_DEF`$ be the change of basis matrix.

[^1]: I prefer to refer to the responses as "positive/negative", instead of
    "correct/incorrect", as this terminology contemplates the possibility of
    non-cognitive as well as cognitive domains.

[^2]: I'm preferring here the notation I use in the draft, that is,
    $`r DIM_INDEX`$ for indexing the dimensions (instead of $`r ITEM_INDEX`$,
    which is also used to index the items), $`r TRAIT_MODULE`$ for the
    coordinates of $`r TRAIT_VECTOR`$ in $`r LS_BASIS`$,
    and $`r TRAIT_VEC_STD_EL`$ for the coordinates in $`r LS_STD_BASIS`$.
    Nevertheless, as Mario reminded me, it is worth noting that this notation
    for the vector coordinates is ambiguous, as we are using the same symbol
    ($`r TRAIT_VECTOR`$) for the "geometric vector" and
    the "algebraic vector" (i.e. column matrix) of its coordinates in
    $`r LATENT_SPACE`$ (this notation is usually accepted though).

>>>>

Let's consider a dichotomous item $`r ITEM_INDEX`$ such that the probability
that respondent with trait $`r LS_BASIS_EQ`$ gives a positive[^1] response is
given by

@@

Let $`r BASIS_CHANGE_DEF`$ be the change of basis matrix; then, if we denote
$`r TRAIT_COORDS_DEF`$ and $`r TRAIT_STD_COORDS_DEF`$ we have that

$$
  `r BASIS_CHANGE_EQ`
$$ {#eq-change-basis}

The $`r DIM_INDEX`$-th column of $`r TRANSFORM_MATRIX`$ is given by the
coordinates of $`r BASIS_VECTOR_ANY`$ in basis $`r LS_STD_BASIS`$. If
$`r BASIS_VEC_STD_COORDS_DEF`$ then[^3]

$$
  `r TRANSFORM_MATRIX_EQ`
$$ {#eq-transform-matrix-def}

In $`r LATENT_SPACE`$ with the standard basis we have the usual scalar product.
In $`r LATENT_SPACE`$ with basis $`r LS_BASIS`$ we define an inner product
$`r INNER_PROD_BASIS`$ such that the norm of $`r TRAIT_VECTOR`$ is invariant.

[^1]: I prefer to refer to the responses as "positive/negative", instead of
    "correct/incorrect", as this terminology contemplates the possibility of
    non-cognitive (as well as cognitive) domains.

[^2]: I'm preferring here the notation I use in the draft, that is,
    $`r DIM_INDEX`$ for indexing the dimensions (instead of $`r ITEM_INDEX`$,
    which is also used to index the items), $`r TRAIT_MODULE`$ for the
    coordinates of $`r TRAIT_VECTOR`$ in $`r LS_BASIS`$, and
    $`r TRAIT_VEC_STD_COMP`$ for the coordinates in $`r LS_STD_BASIS`$.
    Nevertheless, as Mario reminded me, it is worth noting that this notation
    for the vector coordinates is ambiguous, as we are using the same symbol
    ($`r TRAIT_VECTOR`$) for the "geometric vector" and the "algebraic vector"
    (i.e. column matrix) of its coordinates in $`r LATENT_SPACE`$ (this notation
    is usually accepted though).

[^3]: It is ok to use square brackets here? I'm using it instead of parentheses
    because I had already written this (with a function), but if it's not
    correct I change it to parentheses.
